### 2020-11-20 doc/thesis/front/abstract.tex
Add HgbThesisEN template. (#1)

\chapter{Abstract}

This should be a 1-page (maximum) summary of your work in English.