### 2020-11-20 doc/thesis/front/kurzfassung.tex

\chapter{Kurzfassung}

\begin{german}
An dieser Stelle steht eine Zusammenfassung der Arbeit, Umfang
max.\ 1 Seite. 
...
\end{german}