### 2020-12-04 doc/thesis/chapters/fm_radio.tex
start table of contents

%%%%%%%%%%%%%%%%%%%%
\chapter{Signal Processing Theory}
\label{cha:TheThesis}

  \section{Up-Downconversion to/from RF}

  \section{Frequency Modulation}
  see literature/Present\_lec6\_AM\_FM.pdf\\
  see lectures out of HSD/ESD

  \section{Algorithms for Digital FM Demodulation}
    see literature/FmDemodulator.pdf (Sect. 3.3)
    see literature/00476180 Digital FM Demodulator for FM, TV, and Wireless.pdf (Sect. II and III)

    \subsection{Baseband Delay Demodulator}
    \subsection{Phase-Adapter Demodulator}
    \subsection{Phase-Locked Loop (PLL)}
    \subsection{Mixed Demodulator}

%%%%%%%%%%%%%%%%%%%%
\chapter{High-Level Synthesis}
  \section{Introduction / State of the Art}
  \section{Functionality (high level code to HDL)}
  \section{Variants}
    \subsection{C++}
    \subsection{SystemC}
  \section{Advantages / Disadvantages}

%%%%%%%%%%%%%%%%%%%%
\chapter{Implementation}
  \section{Matlab Simulation}
  in fixed point, close to hardware level algorithm

  \section{VHDL}
    \subsection{Channel Selection (IF to Channel-BB)}
    \subsection{Phase Detector}
    \subsection{other Elements}

  \section{HLS}
    \subsection{Channel Selection (IF to Channel-BB)}
    \subsection{Phase Detector}
    \subsection{other Elements}

  \section{Comparison of Metrics}
    \subsection{Implementation effort/time}
    \subsection{Hardware Utilization}
    \subsection{Others}

%%%%%%%%%%%%%%%%%%%%
\chapter{Deployment on Hardware}
  \section{Hardware Platform}
    \subsection{RTL2832u Dongle}
    \subsection{ZedBoard}