### 2020-12-07 doc/thesis/chapters/fm_radio.tex
some more table of content

%%%%%%%%%%%%%%%%%%%%
\chapter{Signal Processing Theory}
\label{cha:TheThesis}

  \section{Modulation from Baseband to RF}
  \section{Modulation from RF to Baseband}

  \section{Frequency Modulation (FM)}
  see literature/Present\_lec6\_AM\_FM.pdf\\
  see lectures out of HSD/ESD
    \subsection{Mathematical description}
    \subsection{Frequency Band, Channel allocation/distribution}

  \section{Algorithms for Digital FM Demodulation}
    see literature/FmDemodulator.pdf (Sect. 3.3)
    see literature/00476180 Digital FM Demodulator for FM, TV, and Wireless.pdf (Sect. II and III)

    \subsection{Baseband Delay Demodulator}
    \subsection{Phase-Adapter Demodulator}
    \subsection{Phase-Locked Loop (PLL)}
    \subsection{Mixed Demodulator}

%%%%%%%%%%%%%%%%%%%%
\chapter{High-Level Synthesis}
  \section{Introduction / State of the Art}
  \section{Functionality (transform high-level code to HDL)}
  \section{Language Support}
    \subsection{C++}
    \subsection{SystemC}
  \section{Coding}
    \subsection{Compiler Directives (\#pragma's)}
    \subsection{Data types}
    \subsection{Functions}
    \subsection{Loops}
    \subsection{Conditional statements}
  \section{Advantages / Disadvantages}
    easy and much faster to get complex HDL code, such as image/video processing, through libraries like OpenCV. However, it is less optimized, and a developer still needs to have a strong hardware background, to be able to understand what is being generated in hardware from x lines of code in HLS/Cpp.

%%%%%%%%%%%%%%%%%%%%
\chapter{Implementation}
  \section{Matlab/Python Model}
  in fixed point, close to hardware level algorithm

  \section{VHDL}
    \subsection{Channel Selection (IF to Channel-BB)}
    \subsection{Phase Detector}
    \subsection{other Elements}

  \section{HLS}
    \subsection{Channel Selection (IF to Channel-BB)}
    \subsection{Phase Detector}
    \subsection{other Elements}

  \section{Comparison of Metrics}
    \subsection{Implementation effort/time}
    \subsection{Hardware Utilization}
    \subsection{Others}

  \section{Common Testbench}
    \subsection{Architecture (same tb for VHDL and HLS-generated HDL)}
    \subsection{Framework cocotb, with ghdl compiler}
        Instantiate both HDL models in the testbench.\\
        Display a direct comparison of outputs in graphs. This is practical, since the cocotb framework runs in python and graphs can be generated using Python's matplotlib.

  \section{Develop HLS compiler independent code (optional)}
    \subsection{Challenges}
     Different compilers use different \#pragmas, etc.\\
     The \#pragmas need to be within the code, at the appropriate positions, which makes it difficult to write compiler independent code....

%%%%%%%%%%%%%%%%%%%%
\chapter{Deployment on Hardware}
  \section{Hardware Platform}
    \subsection{RTL2832u Dongle}
    \subsection{ZedBoard}