### 2021-07-02 doc/thesis/chapters/system-architecture-and-concept.tex
write stuff

%%%%%%%%%%%%%%%%%%%%
\chapter{System Architecture/Concept}
\label{cha:SystemArchitecture/Concept}

  \section{Overview}

-  implementation of a model and DSP concept in Matlab \\
-  implementation in VHDL and HLS \\
-  both use the same testbench source \\
-  both are integrated into FPGA
-  the FPGA output is read back into the firmware, for analysis \\

  \section{Block Diagram (with details)}
    \subsection{describe blocks}

  \section{Test Environment}