### 2021-07-15 doc/thesis/chapters/deployment-on-hardware.tex
bd_impl_vivado diagram

%%%%%%%%%%%%%%%%%%%%
\chapter{Deployment on Hardware}
\label{cha:DeploymentOnHardware}

%%%%%%%%%%%%%%%%%%%%
\section{GNU Radio Transmitter}

%%%%%%%%%%%%%%%%%%%%
\section{Hardware Platform}
ZedBoard

%%%%%%%%%%%%%%%%%%%%
\subsection{RTL-SDR}
RTL2832U USB Dongle

%%%%%%%%%%%%%%%%%%%%
\section{Block diagram}

\includepicture [1.0] {Block diagram of the Vivado implementation.} {bd_impl_vivado} {img/draw.io/bd_impl_vivado}