### 2021-07-21 doc/thesis/chapters/conclusion-and-prospect.tex
hls

%%%%%%%%%%%%%%%%%%%%
\chapter{Conclusion and Prospect}
\label{cha:ConclusionAndProspect}

% NOTES:
%  - use Matlab Simulink HDL Coder and System Generator
%    as additional implementation variants
%  - explore HLS optimizations
%  - explore HLS libraries (e.g. DSP library --> pre-made FIR filter)
%  - create another IP in the same system
%     -- due to the modular system design this is easy to do
%
%
%
%
%