### 2021-07-22 doc/thesis/chapters/conclusion-and-prospect.tex
product design

%%%%%%%%%%%%%%%%%%%%
\chapter{Conclusion and Prospect}
\label{cha:ConclusionAndProspect}

% NOTES:
%  - use Matlab Simulink HDL Coder and System Generator
%    as additional implementation variants
%  - use Intel oneAPI HLS compiler
%     -- try to develop compiler-independent code
%     -- how platform-independent is HLS?
%  - explore HLS optimizations
%  - explore HLS libraries (e.g. DSP library --> pre-made FIR filter)
%  - create another IP in the same system
%     -- due to the modular system design this is easy to do
%  - use Linux OS on the SoC, to enable live-data streaming with the RTL-SDR
%
%
%
%
%