### 2021-08-22 doc/thesis/back/supplementary-materials.tex
Thesis chapters text. (#30)

\chapter{Supplementary Materials}
\label{app:materials}

A list of supplementary data, that is included with this thesis.

\section{PDF Files}
\begin{FileList}{/}
\fitem{thesis.pdf} Master thesis
\end{FileList}

\section{FM Radio Receiver Project}
\begin{FileList}{/fm-radio}
\fitem{/doc} Documentation, Analysis tools and this thesis document.
\fitem{/hardware} Hardware development, HLS and VHDL.
\fitem{/sim} Simulation, including GNU Radio and Matlab.
\end{FileList}

\vspace{5 mm}
\noindent
The FM Radio Receiver project can also be found on GitHub:\\
\url{https://github.com/wurmmi/fm-radio}